---
title: "data_analysis"
author: "Miles Eng"
date: "11/12/2023"
output: html_document
---
```{r load-packages}

library(tidyverse)
library(tidymodels)
library(lubridate)
library(patchwork)

```

Drivers data is useless since it reads from a queue that doesn't have all drivers (its fine though we have all the same data in rides)

``` {r load-data}

t1rides_1 <- read_csv('../results/t1rides_trial1.csv') |>
  mutate(trial = 1)
t1rides_2 <- read_csv('../results/t1rides_trial2.csv') |>
  mutate(trial = 2)
t1rides_3 <- read_csv('../results/t1rides_trial3.csv') |>
  mutate(trial = 3)
t1rides_4 <- read_csv('../results/t1rides_trial4.csv') |>
  mutate(trial = 4)
t1rides_5 <- read_csv('../results/t1rides_trial5.csv') |>
  mutate(trial = 5)

t1rides <- do.call("rbind", list(t1rides_1, t1rides_2, t1rides_3, t1rides_4, t1rides_5))

t2rides <- read_csv('../results/t2rides.csv')
t3rides <- read_csv('../results/t3rides.csv')
t4rides <- read_csv('../results/t4rides.csv')
# t5rides <- read_csv('../results/t5rides.csv')

t1rides <- t1rides |>
  select(-drivers_in_queue) |>
  mutate(task = 1)

t2rides <- t2rides |>
  mutate(task = 2)

t3rides <- t3rides |>
  select(-dest_lat, -dest_lon, -source_lat, -source_lon, -driver_lat, -driver_lon) |>
  mutate(task = 3,
         match_runtime_sec = 0)

t4rides <- t4rides |>
  mutate(task = 4) 

# t5rides <- t5rides |>
#   mutate(task = 5)

rides <- do.call("rbind", list(t1rides, t2rides, t3rides, t4rides))

rides <- rides |>
  mutate(driver_to_passenger_hours = driver_to_passenger,
         passenger_to_dest_hours = passenger_to_dest) |>
  select(-driver_to_passenger, -passenger_to_dest)

rides <- rides |>
  mutate(time_to_match_sec = as.numeric(ride_match_datetime - passenger_start_datetime),
         passenger_wait_time_sec = time_to_match_sec + (driver_to_passenger_hours * 60 * 60),
         passenger_wait_time_min = passenger_wait_time_sec / 60,
         driver_to_passenger_min = driver_to_passenger_hours * 60,
         passenger_to_dest_min = passenger_to_dest_hours * 60,
         driver_profit_hours = passenger_to_dest_hours - driver_to_passenger_hours,
         driver_profit_min = driver_profit_hours * 60,
         did_driver_profit = if_else(driver_profit_hours > 0, 1, 0),
         wait_category = case_when(passenger_wait_time_min < 5 ~ "short",
                                   passenger_wait_time_min < 20 ~ "medium",
                                   .default = "long"),
         task = as.factor(task),
         wait_category = fct_relevel(wait_category, "short", "medium", "long"),
         total_d1_time_min = as.numeric(ride_match_datetime - passenger_start_datetime) / 60 + driver_to_passenger_min + passenger_to_dest_min)

```

# Exploratory Data Visualization

```{r performance}

t1runtimes <- data.frame(
  task = c(1,1,1,1,1),
  total_runtime_sec = c(1193.0641, 1130.689230203, NA, NA, NA),
  queue_runtime_sec = c(80.5091679, 224.383005142211, NA, NA, NA)
)

t2runtimes <- data.frame(
  task = c(2,2,2,2,2),
  total_runtime_sec = c(1322.442229, 1148.78903698, NA, NA, NA),
  queue_runtime_sec = c(82.2850890, 208.86551904, NA, NA, NA)
)

t3runtimes <- data.frame(
  task = c(3,3,3,3,3),
  total_runtime_sec = c(13892.306724071, NA, NA, NA, NA),
  queue_runtime_sec = c(70.143744, NA, NA, NA, NA)
)

t4runtimes <- data.frame(
  task = c(4,4,4,4,4),
  total_runtime_sec = c(1242.4165439605, NA, NA, NA, NA),
  queue_runtime_sec = c(0.772353887557, NA, NA, NA, NA)
)

t5runtimes <- data.frame(
  task = c(5,5,5,5,5),
  total_runtime_sec = c(NA, NA, NA, NA, NA),
  queue_runtime_sec = c(NA, NA, NA, NA, NA)
)

runtime_data = do.call("rbind", list(t1runtimes, t2runtimes, t3runtimes, t4runtimes, t5runtimes))

```

D1. Passengers want to be dropped off as soon as possible, that is, to minimize 
the amount of time (in minutes) between when they appear as an unmatched 
passenger, and when they are dropped off at their destination.

```{r}

rides |>
  ggplot(aes(x = passenger_start_datetime, y = total_d1_time_min, color = task)) +
  geom_point() +
  labs(title = "Fig 1. Total Passenger Wait Time",
       x = "Passenger Ride Request Date/Time",
       y = "Time from Ride Request to Completion (min)",
       color = "Task") +
  theme_bw()

rides |>
  ggplot(aes(x = total_d1_time_min, fill = task)) +
  geom_boxplot()

```


D2. Drivers want to maximize ride profit, defined as the number of minutes they 
spend driving passengers from pickup to drop-off locations minus the number 
of minutes they spend driving to pick up passengers. For simplicity, there is 
no penalty for the time when a driver is idle.

```{r}

rides |>
  ggplot(aes(x = task, fill = as.factor(did_driver_profit))) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Fig 2. Driver Ride Profit",
       x = "Driver Profit",
       fill = "Did Driver Profit") +
  scale_fill_discrete(labels = c("False", "True"))

```

D3. Your algorithms should be empirically efficient and scalable.

```{r}

runtime_data |>
  filter(!is.na(total_runtime_sec)) |>
  mutate(task = as.factor(task)) |>
  group_by(task) |>
  summarize(mean_total_runtime_sec = mean(total_runtime_sec)) |>
  ggplot(aes(x = task, y = mean_total_runtime_sec, fill = task)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = signif(round(mean_total_runtime_sec, digits = 3))), nudge_y = 500) +
  labs(title = "Total Runtime of NotUber Matching Algorithm",
       x = "Task",
       y = "Total Runtime (sec)") +
  scale_fill_manual(values=c("#1d4941", "#305c54", "#42665f", "#6e9991", "#93c0b8")) +
  theme_bw() +
  theme(legend.position = "none")

runtime_data |>
  filter(!is.na(queue_runtime_sec)) |>
  mutate(task = as.factor(task)) |>
  group_by(task) |>
  summarize(mean_queue_runtime_sec = mean(queue_runtime_sec)) |>
  ggplot(aes(x = task, y = mean_queue_runtime_sec, fill = task)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = signif(round(mean_queue_runtime_sec, digits = 3))), nudge_y = 5) +
  labs(title = "Runtime of Passenger and Driver Queue Construction",
       x = "Task",
       y = "Queue Runtime (sec)") +
  scale_fill_manual(values=c("#1d4941", "#305c54", "#42665f", "#6e9991", "#93c0b8")) +
  theme_bw() +
  theme(legend.position = "none")

```


